Rust for Scientific Computing

### 2025-09-15 paper/main.tex
continue writing introduction

Importance of programming in science
\begin{enumerate}
    \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
    \item Cite some groundbreaking examples which were possible due to programming in
        science (maybe weather)
    \item
\end{enumerate}

Many languages have been devised to target specific realms which have emerged over the past decades.
They span across a range of abstraction levels, ranging from low-level languages to high-level ones.
The low-level operations level mostly deals with instructions on a level close to the actual
hardware such as memory allocation, pointer arithmetic or by using assembly instructions directly.
Meanwhile, high-level operations usually combine many lower-level ones to perform a single task.
An example for a high-level operation would be the addition of two matrices which might allocate the
required memory, fill it according to the given values and then call the appropriate math function
to pairwise add their entries.
Languages which can operate effectively on a lower level often bear the most promise for maximum
performance which is crucial in any scientific application.
These languages are mostly compiled using an optimizing compiler which enables many of their
performance achievements.

>>>>

Many languages are able to solve many different problems in a general-purpose fashion but some have
also been developed to target specific realms which have emerged over the past decades.
Early achievements of the first commercially available programming language FORTRAN are still used
today in the Modular Ocean Model~\cite{Adcroft2019,Griffies2012}.
The development of the C programming language in 1972 enabled milestones such as the Unix operating
system~ \cite{Ritchie1978} which layed the foundation for many major scientific achievements.
With the advent of C++ in 1985 further contributions were made such as the root project which
is a "Object Oriented Data-Analysis Framework"~\cite{Brun1997} used by the CERN
project~\cite{https://doi.org/10.7483/opendata.atlas.l806.5cku}.
In more recent years  deep learning tools such as pytorch~\cite{Ansel_PyTorch_2_Faster_2024},
tensorflow~\cite{tensorflow2015-whitepaper} have seen widespread use which are programmed on a
higher abstraction level using Python.
The latter enabled the development of AlphaFold which is provides AI-based protein structure
prediction~\cite{Jumper2021}.

% Importance of programming in science
% \begin{enumerate}
%     \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
%     \item Cite some groundbreaking examples which were possible due to programming in
%         science (maybe weather)
%     \item
% \end{enumerate}

As the preceding exmamples show, even a single programming language can span across multiple
abstraction levels, ranging from low-levels of operation to high-level ones.
The former mostly deal with instructions on a level close to the actual hardware such as memory
allocation, pointer arithmetic or by using assembly instructions directly.
Meanwhile, high-level operations usually combine many lower-level ones to perform a single task.
An example for a high-level operation would be the addition of two matrices which might allocate the
required memory, fill it according to the given values and then call the appropriate math function
to pairwise add their entries.
Languages which can operate effectively on a lower level often bear the most promise for maximum
performance which is crucial in any scientific application.
Of these languages almost all are compiled using an optimizing compiler which enables many of their
performance achievements.
